---
title: "BCB546_Project_Peng_et_al_2015_Figure_3_Heat_maps"
author: "Tyler L Foster"
date: "4/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This R markdown file is written and documented by Tyler Foster. The aim of this coding file is to convert gene expression vales from Peng_et_al_2015 found in "Supplemental file 5.xls" to create a heat map for each of the following species: Arabidopsis, Wheat and Barley.


# Dependencies
```{r}
if (!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)

if (!require("gplots")) install.packages("gplots")
library(gplots)

if (!require("readxl")) install.packages("readxl")
library(readxl)
```
###################################### Arabidopsis ######################################

# Data Wrangling
```{r}
#read in the 97-2003 file, change the sheet accordingly, skip the nasty 1st row, sheet 1 is arabidopsis 
arabidopsis_expression_v <- read_excel("../Peng_Data/Supplemental file 5.xls", sheet = 1, skip = 1)

#change 1st column name to 'Gene'
colnames(arabidopsis_expression_v)[1] <- "Gene"

#Capitalize the first character
arabidopsis_expression_v <- arabidopsis_expression_v %>% rename_with(str_to_title)

#creates a datamatrix, which heatmap.2 can utilize later, to generate the desired plot
arabidopsis_matrix <- data.matrix(arabidopsis_expression_v)

#change the row names in the matrix to the gene names from our dataframe
row.names(arabidopsis_matrix) <- arabidopsis_expression_v$Gene

#selecting all rows and columns 2-64 of our datamatrix as column 1 is a gene number
arabidopsis_matrix <- arabidopsis_matrix[,2:64]

#Creates a color palette for our heatmap
colfunc <- colorRampPalette(c("firebrick", "mediumblue", "green"))
```

# Re-creation of Original Arabidopsis Heat Map
```{r}
bmp(file="../Results/Fig_3/Original_Figure3a.bmp",
   width=16, height = 10, units = "in", res=250)
heatmap.2(arabidopsis_matrix,
          dendrogram = "col", #only draw a column dendogram
          breaks = 100, #determines the number of color transitions
          trace = "none", #turns off the trace lines inside the heat map
          margins = c(5,0), #sets margins around plot
          labRow = FALSE, #turns off row labels
          col = colfunc(99), #use the color palette we created above
          cexCol = 0.7, #changes column label font
          lmat = rbind(c(0,3,0),c(2,1,0),c(0,4,0)), #manually placing key objects of figure
          lwid = c(0.05, 0.1, 0.05), #changes dimensions of the array cell widths
          lhei = c(0.43, 2.5, 0.6), #alter dimensions of array cell heights
          key.title = NA, #turns off key title
          density.info="none", #turns off the trace lines in the key
          key.par=list(mar=c(3.2, 0, 4,0) ), #tweaks specific key parameters
)
```

# Improved Arabidopsis Heat Map
```{r}
#Creates a 2nd color palette for our heatmap
colfunc2 <- colorRampPalette(c("blue2", "white", "red2"))

bmp(file="../Results/Fig_3/Ice_Cream_Bean's_Figure3a.bmp",
   width=16, height = 10, units = "in", res=250)
heatmap.2(arabidopsis_matrix,
          dendrogram = "col", #only draw a column dendogram
          breaks = 100, #determines the number of color transitions
          trace = "none", #turns off the trace lines inside the heat map
          margins = c(5,0), #sets margins around plot
          labRow = FALSE, #turns off row labels
          col = colfunc2(99), #use the color palette we created above
          cexCol = 0.7, #changes column label font
          lmat = rbind(c(0,3,0),c(2,1,0),c(0,4,0)), #manually placing key objects of figure
          lwid = c(0.05, 0.1, 0.05), #changes dimensions of the array cell widths
          lhei = c(0.43, 2.5, 0.6), #alter dimensions of array cell heights
          key.title = NA, #turns off key title
          density.info="none", #turns off the trace lines in the key
          key.par=list(mar=c(3.2, 0, 4,0) ), #tweaks specific key parameters
          key.xtickfun = function() {breaks = pretty(parent.frame()$breaks)          
            breaks = breaks[c(1,length(breaks))]             
            list(at = parent.frame()$scale01(breaks), labels = breaks) #changes the tick marks on key
            })
```

###################################### Wheat ######################################

# Data Wrangling
```{r}
#read in the 97-2003 file, change the sheet accordingly, skip the nasty 1st row, sheet 2 is wheat 
wheat_expression_v <- read_excel("../Peng_Data/Supplemental file 5.xls", sheet = 2, skip = 1)

#change 1st column name to 'Gene'
colnames(wheat_expression_v)[1] <- "Gene"

#Capitalize the first character
wheat_expression_v <- wheat_expression_v %>% rename_with(str_to_title)

#creates a datamatrix, which heatmap.2 can utilize later, to generate the desired plot
wheat_matrix <- data.matrix(wheat_expression_v)

#change the row names in the matrix to the gene names from our dataframe
row.names(wheat_matrix) <- wheat_expression_v$Gene

#selecting all rows and columns 2-14 of our datamatrix as column 1 is a gene number
wheat_matrix <- wheat_matrix[,2:14]
```

# Re-creation of Original Wheat Heat Map
```{r}
bmp(file="../Results/Fig_3/Original_Figure3b.bmp",
    width=6, height = 4, units = "in", res=250)
heatmap.2(wheat_matrix,
          dendrogram = "col", #only draw a column dendogram
          breaks = 100, #determines the number of color transitions
          trace = "none", #turns off the trace lines inside the heat map
          margins = c(10,0), #widens margins around plot
          labRow = FALSE, #turns off row labels
          col = colfunc(99), #use the color palette we created above
          cexCol = 0.9, #changes column label font
          lmat = rbind(c(0,3,0),c(2,1,0),c(0,4,0)), #manually placing key objects of figure
          lwid = c(0.05, 0.05, 0.05), #changes dimensions of the array cell widths
          lhei = c(0.23, 1.5, 0.6), #alter dimensions of array cell heights
          key.title = NA, #turns off key title
          density.info="none", #turns off the trace lines in the key
          key.par=list(mar=c(3.2, 0, 2.5,0) ), #tweaks specific key parameters
)
```

# Improved Wheat Heat Map
```{r}
bmp(file="../Results/Fig_3/Ice_Cream_Bean's_Figure3b.bmp",
    width=6, height = 4, units = "in", res=250)
heatmap.2(wheat_matrix,
          dendrogram = "col", #only draw a column dendogram
          breaks = 100, #determines the number of color transitions
          trace = "none", #turns off the trace lines inside the heat map
          margins = c(10,0), #widens margins around plot
          labRow = FALSE, #turns off row labels
          col = colfunc2(99), #use the color palette we created above
          cexCol = 0.9, #changes column label font
          lmat = rbind(c(0,3,0),c(2,1,0),c(0,4,0)), #manually placing key objects of figure
          lwid = c(0.05, 0.05, 0.05), #changes dimensions of the array cell widths
          lhei = c(0.23, 1.5, 0.6), #alter dimensions of array cell heights
          key.title = NA, #turns off key title
          density.info="none", #turns off the trace lines in the key
          key.par=list(mar=c(3.2, 0, 2.5,0) ), #tweaks specific key parameters
          key.xtickfun = function() {breaks = pretty(parent.frame()$breaks)          
            breaks = breaks[c(1,length(breaks))]             
            list(at = parent.frame()$scale01(breaks), labels = breaks) #changes the tick marks on key
            })
```

###################################### Barley ######################################

# Data Wrangling
```{r}
#read in the 97-2003 file, change the sheet accordingly, skip the nasty 1st row, sheet 3 is barley 
barley_expression_v <- read_excel("../Peng_Data/Supplemental file 5.xls", sheet = 3, skip = 1)

#change 1st column name to 'Gene'
colnames(barley_expression_v)[1] <- "Gene"

#Capitalize the first character
barley_expression_v <- barley_expression_v %>% rename_with(str_to_title)

#creates a datamatrix, which heatmap.2 can utilize later, to generate the desired plot
barley_matrix <- data.matrix(barley_expression_v)

#change the row names in the matrix to the gene names from our dataframe
row.names(barley_matrix) <- barley_expression_v$Gene

#selecting all rows and columns 2-16 of our datamatrix as column 1 is just a gene number
barley_matrix <- barley_matrix[,2:16]
```

# Re-creation of Original Barley Heat Map
```{r}
bmp(file="../Results/Fig_3/Original_Figure3c.bmp",
    width=6, height = 4, units = "in", res=250)
heatmap.2(barley_matrix,
          dendrogram = "col", #only draw a column dendogram
          breaks = 100, #determines the number of color transitions
          trace = "none", #turns off the trace lines inside the heat map
          margins = c(10,0), #widens margins around plot
          labRow = FALSE, #turns off row labels
          col = colfunc(99), #use the color palette we created above
          cexCol = 0.9, #changes column label font
          key.title = NA, #turns off key title
          density.info="none", #turns off the trace lines in the key          
          lmat = rbind(c(0,3,0),c(2,1,0),c(0,4,0)), #manually placing key objects of figure
          lwid = c(0.05, 0.05, 0.05), #changes dimensions of the array cell widths
          lhei = c(0.23, 1.5, 0.6), #alter dimensions of array cell heights
          key.par=list(mar=c(3.2, 0, 2.5,0) ), #tweaks specific key parameters
)
```

# Improved Barley Heat Map
```{r}
bmp(file="../Results/Fig_3/Ice_Cream_Bean's_Figure3c.bmp",
    width=6, height = 4, units = "in", res=250)
heatmap.2(barley_matrix,
          dendrogram = "col", #only draw a column dendogram
          breaks = 100, #determines the number of color transitions
          trace = "none", #turns off the trace lines inside the heat map
          margins = c(10,0), #widens margins around plot
          labRow = FALSE, #turns off row labels
          col = colfunc2(99), #use the color palette we created above
          cexCol = 0.9, #changes column label font
          key.title = NA, #turns off key title
          density.info="none", #turns off the trace lines in the key          
          lmat = rbind(c(0,3,0),c(2,1,0),c(0,4,0)), #manually placing key objects of figure
          lwid = c(0.05, 0.05, 0.05), #changes dimensions of the array cell widths
          lhei = c(0.23, 1.5, 0.6), #alter dimensions of array cell heights
          key.par=list(mar=c(3.2, 0, 2.5,0) ), #tweaks specific key parameters
          key.xtickfun = function() {breaks = pretty(parent.frame()$breaks)          
            breaks = breaks[c(1,length(breaks))]             
            list(at = parent.frame()$scale01(breaks), labels = breaks) #changes the tick marks on key
            })
```